---
title: "Predicting Firm Location Decisions"
author: "John Ohakim"
date: "July 18, 2018"
output: html_document
---

 The conventional statistical and econometric techniques, such as regression analysis, that are employed by economists often work well. However, given the issues that are unique to very large datasets, economic analysis may require the use of entirely different statistical tools. 

  Varian (2014) identifies three of these issues: First, the size of the data involved may require more powerful data manipulation tools, on which linear regressions perform poorly. Second, there may be more potential predictors than appropriate for estimation. Hence, the need to do some kind of variable selection arises which is outside our domain knowledge. Third, large datasets may allow for more flexible relationships than simple linear models. Machine learning (ML) and deep learning algorithms (DL) such as decision trees, Support Vector Machine (SVM), Random Forests, and Recurrent Neural Network (RNN), may allow for more effective ways to model complex relationships. The goal of this project is to employ economic data using machine learning and deep learning algorithms to answer a familiar economic question. 

  Although the literature on Foreign Direct Investment (FDI) is vast, relatively less attention has been devoted to empirically analyzing the emergence of FDI from developing countries to other developing countries and the resulting effects of these flows in host and home countries. The emergence, and growth, of the operations of these multinational corporations from the South may have important implications for developing countries for a number of reasons. We seek to predict the location decisions, in developing country economies, of emerging market multinational firms and their developed country counterparts. 

  To do this, we employ a novel times series data set and RNN to see if it can predict the location decisions of firms. First, we gather the data fDi Markets data set of the Financial Times that tracks cross-border, greenfield investments across over two hundred countries worldwide, with real-time monitoring of investment projects, capital investment and job creation. Using the caret and keras packages, we split our dataset into a training and test set; the training set is used to train the model, and the test used to predict our outcome variable. This approach may provide a model that avoids overfitting and provides external, something that previous economic models fail to address in a convincing way. Finally, We compare the performance of this model with the baseline regression model.
  
  Data and Sources
  
  This dataset is constructed from four different data sources:
1. fDi Markets of the Financial Times (https://www.fdimarkets.com/)
A unique dataset that tracks cross-border, greenfield and brownfield investmemts across 200 countries worldwide, with realtime monitoring of investment projects, capital expenditure, and job creation.

2. The Geodist database (http://www.cepii.fr/CEPII/en/bdd_modele/presentation.asp?id=30)
This provides country-specific spatial features for over 225 countries.

3. Total Economy database (https://www.conference-board.org/data/economydatabase/index.cfm?id=25667)
This contains annual time series data on Gross Domestic Product, population, employment, labor productivity, and so on. 

4. Worldwide Governance Indicators (http://info.worldbank.org/governance/wgi/index.aspx#home)
This reports governance indicators which proxy for the quality of institutions in a country along 6 dimensions. 

  Together, this dataset allows for the identification of the origin and destination of cross-border investments. This granularity permits the construction of a novel bilateral times series dataset, with over 30 variables/features over a 13 year period. 
  
  
```{r}
gravity <- read.csv("C://Users//johno//Dropbox//Econ Grad//Dissertation//Proposal//South-South//Paper_2//Data//Gravity.csv")

knitr::kable(
  head(gravity),
  caption = "Glimpse of the Data Set"
)

names(gravity)
str(gravity)
summary(gravity)

library(tidyverse)

gravity$id <- with(gravity, ave(rep(1, nrow(gravity)), Year, FUN = seq_along))

gravity <- gravity %>% select(id, Host_Country:NR, starts_with('id'))



```



Data Visualizations of Receiving/Host Economies

```{r}

ggplot(data = gravity) +
  geom_line(mapping = aes(x = Year, y = FDI)) +
  facet_wrap(~ Host_Country) +
  ggtitle("FDI in Emerging Markets (2003-2015)")

ggplot(data = gravity, mapping = aes(x = log(FDI), y = log(mktsize))) +
  geom_point(mapping = aes(color = continent)) +
  geom_smooth(
    #data = filter(gravity, continent == "Asia"),
    se = F
  ) +
  ggtitle("FDI vs Size of the Market (2003-2015)")

ggplot(data = gravity) +
  stat_count(mapping = aes(x = Direction, fill = continent)) +
  ggtitle("Direction of FDI Flows (North-South vs South-South)")
  
ggplot(data = gravity) +
  geom_bar(
  mapping = aes(x = Direction, y = ..prop.., group = 1)) +
  ggtitle("Direction of FDI Flows (North-South vs South-South)")

```
